### 2019-08-23 Manuscripts/post-cogpsych/post-cogpsych.tex
Updated section 7.2 of cog sci manuscript

Another factor we examine with this category structure is the effect of boundaries. The represenativeness model prefers to generate categories located
at the boundaries of the stimulus features. If at least one of these boundaries
is removed by utilizing a toroidal feature (one with no boundaries), such as the rotation of a line. The represenativeness model will strongly prefer to generate categories on the non-toroidal or bounded feature as it still has a boundary. Experiment 3 manipulates the stimulus space. One condition uses the same stimuli as before, squares varying in lightness and size, both of which are bounded features. The other condition use a new set of stimuli:
circles that vary in size (a bounded feature) and orientation (an
unbounded feature; Figure \ref{fig:e3-circles}a). Before describing the
details of the experiment, we examine each model's
predictions with stimuli that comprise one bounded and one unbounded feature.

\begin{figure}[p]
    \begin{center} 
      \includegraphics[width=.8\textwidth]{figs/e3-circles.pdf}
      \caption{Circle stimulus domain (a) and predictions from contrast models
        PACKER (b) and the representativeness model (c). The unbounded feature
        (orientation) is represented on the y-axis, while the bounded feature
        (size) is represented on a x-axis. As with the square stimuli, feature
        and direction assignment were counterbalanced across participants (in
        the displayed example, circles are rotated clockwise up the y-axis.) }
      \label{fig:e3-circles}
    \end{center}
\end{figure}

@@

The single unbounded feature prevents ... While the representativeness model
more closely resembles PACKER under this condition, note that the former still
more strongly predicts the generation of the first Betas exemplars at the
bounded edges compared to PACKER (Figure \ref{fig:e3-circles}c.) 
% For the representativeness model, the single unbounded feature essentially
% prevents the formation of edge effects along that feature. Consequently, this results in higher probabilities of the generating the first Beta exemplars along the edges of the bounded feature (Figure \ref{fig:e3-circles}c). As this
% model does not explicitly recognize similarities as distances along the feature
% space (unlike PACKER), it does not experience the compression in density regions
% along the unbounded feature axis. Intuitively, this suggests that the
% representativeness model (which prefers generating exemplars at the boundaries) will prefer to generate stimuli on the boundaries of the bounded feature with some amount of variance over unbounded feature.

@@

We recruited 89 participants from Amazon Mechanical Turk who were randomly
assigned to the bounded, square stimuli ($N=46$), or to the toroidal, circle
stimuli ($N=43$). The square stimuli were exactly the same as in Experiments 1
and 2. The circle stimuli were made up of unfilled circles with a radial line
extending from its center to its edge. These stimuli varied along two features
-- their diameter ranging from 3.0--5.8 cm, and their orientation ranging from
$2.587$--$2\pi + 2.587$ radians, with $0$ radians indicating a completely
horizontal line. The line widths of the circles and radial line were set to 0.05
cm. Similar to Experiments 1 and 2, participants first observed a training phase
where four Alpha exemplars were presented across three blocks followed by the
generation phase. The training condition was the Corner category for all
participants.

@@

%Repeat hypotheses

@@

Participants produced circle stimuli with significantly lower Y-range (lower for the unbounded feature) as compared with the square stimuli ($t(87) = 4.94$, $p = .001$).  This
provides preliminary support for PACKER, which predicts generating exemplars
within a shorter range along the unbounded feature compared to a bounded
feature. Otherwise, there were no significant differences between the distributional statistics of each
generated category between square and circle stimuli (Figure
\ref{fig:e3-statsboxes})

The contrast models were fit by optimizing their fit to all experiments in this
article (Experiments 1 through 3). Fit statistics and parameter values are
presented in Table \ref{table:corner-model-fits}. Although not our main focus in
this section, we also provide the statistics for the classical models --
unsurprisingly, they do not perform as well as the contrast models. Overall,
the representativeness model emerges as the best-fitting model in aggregate, although by a much smaller amount than Experiments 1 and 2. However, analyzing data in aggregate can be misleading.

%by a much smaller lead compared to fits to only Experiments 1 and 2. At an individual level, 67\% of participants in Experiment 3 were better
%fit by PACKER than the representativeness model.

Table \ref{table:corner_counts} presents frequencies of the best-fitting model for participants in each condition. Overall the representativeness model was significantly worse at capturing the results of Experiment 3 (it fit 29 out of the 89 participants better than PACKER; two-tailed Binomial test $p < 0.005$). Further, this is mostly due to representativeness model capturing participants in the unbounded circle condition worse than  PACKER (11 of 43 better fit by representativeness; two-tailed Binomial test, $p<0.005$). The representativeness model only trended towards fitting participants in the bounded, Square condition better (18 out of 46; two-tailed Binomial test $p = 0.18$). Further, 35\% of participants in the bounded condition generated their first Beta exemplars along the
edges of the feature space, but only 12\% of participants in the toroidal, circle condition generated their first Beta exemplars along the edges of
the feature space. These proportions more closely match the predictions of
PACKER (33\% probability of generating first Beta exemplars at the edges for
squares, and 24\% probability at edges for circles) compared to the
representativeness model (53\% of the former, 49\% probability of the latter). Thus, the representativeness model's success at capturing Experiments 1 and 2 is in part due to the nature of the stimulus features, as well as the structure of contrasting category.
%In addition, these results agree with our prediction that the representativeness model would be poorer in predicting data from circle stimuli (26\% of participants observing circle stimuli) compared to square stimuli (39\% participants observing square stimuli).

\begin{table} \centering
  \caption{Frequencies of individual participants best fit by each contrast
    model  broken down by stimulus type.}
\label{table:corner_counts}
\begin{tabular}{ l l l l l} \\ Stimulus Type & \textbf{PACKER} &
                                                               \textbf{Representativeness}& Total & Two-tailed Binomial test $p$-value \\
  \hline 
  Squares & 28 & 18 & 46 & $p < 0.005$\\
  Circles & 32 & 11 & 43 & $p \approx 0.18$\\
  Total   & 60 & 29 & 89 & $p < 0.005$\\
\end{tabular}
\end{table}
\begin{table} \centering
  \caption{Results of model-fitting to the combined datasets from Experiments 1,
    2 and 3. Smaller AIC values correspond to better model fits (adjusted
    for number of parameters.)}
\label{table:corner-model-fits}

\begin{tabular}{ l l l l} \\ \\ \textbf{PACKER} & \textbf{Copy \& Tweak} &
\textbf{Representativeness} & \textbf{Hierarchical Bayesian} \\ 
  \hline $AIC = 12023$ &$AIC = 12779$ &$AIC = 11982$  &$AIC = 12976$\\ 
  $L = -6008$          &$L = -6388$   &$L = -5987$    &$L = -6484$\\ 
  $c = 0.51$           &$c = 3.24$    &$\kappa = 6.73$&$\kappa = 0.00$  \\ 
  $\theta_c = 3.03$    &$\theta_c = 0$ (fixed)  &$\nu = 1.00$ &$\nu = 3.27$\\
  $\theta_t = 3.46$    &$\theta_t = 2.90$       &$\lambda = 1.90$ &$\lambda = 0.07$\\
                       &              &$\theta = 2.45$ &$\theta = 3.15$\\
\end{tabular}
\end{table}

%Ultimately, these results suggest that PACKER's exemplar dissimilarity account
%of contrast, as opposed to the representativeness account, is more appropriate
%for the Corner category type. On almost every measure of analysis in this
%section, PACKER outperformed the representativeness model. Although the
%representativeness model best fit the data from Experiments 1 and 2, it
%struggled when the learned category occupied the extreme corners of the stimulus
%space -- both in cases where the stimuli possess bounded and unbounded features.
%
%[Work on transition time permitting]

@@

Nonetheless, these classic effects are a core element of the phenomenology of
category generation, and PACKER does not include any mechanisms that explain
them. Instead, through the development and evaluation of the PACKER model, we
have sought to add new elements into such a phenomenology: The broad and strong
influence of category contrast, and the interdependence between category
location and distributional structure. It may be possible to combine the
hierarchical Bayesian approach proposed by \cite{jern2013probabilistic} with
PACKER's underlying claims to obtain a ``best of both worlds'' model, capable of
explaining the role of contrast in category generation, as well as the emulation
of distributional structure. However, as noted in the introduction, the
incorporation of category contrast is antithetical to the core principles of a
traditional, semi-conjugate Bayesian approach. This suggests that category
generation is a fundamentally different computational-level problem
\citep[different from those posed by][]{jern2013probabilistic,kemp2014taxonomy}, which is supported by the represenativeness form of their model performing much better.

>>>>

Another factor we examine with this category structure is the effect of
boundaries. The representativeness model prefers to generate categories located
at the boundaries of the stimulus features. If at least one of these boundaries
is removed by utilizing a toroidal feature (one with no boundaries), such as the
rotation of a line.[Note to Xian: Something weird going on here, remember to edit.] The representativeness model will strongly prefer to
generate categories on the non-toroidal or bounded feature as it still has a
boundary. Experiment 3 manipulates the stimulus space. One condition uses the
same stimuli as before, squares varying in lightness and size, both of which are
bounded features. The other condition use a new set of stimuli: circles that
vary in size (a bounded feature) and orientation (an unbounded feature; Figure
\ref{fig:e3-circles}a). Before describing the details of the experiment, we
examine each model's predictions with stimuli that comprise one bounded and one
unbounded feature.

\begin{figure}[p]
    \begin{center} 
      \includegraphics[width=.8\textwidth]{figs/e3-circles.png}
      \caption{Circle stimulus domain (a) and predictions from contrast models
        PACKER (b) and the representativeness model (c). The unbounded feature
        (orientation) is represented on the y-axis, while the bounded feature
        (size) is represented on a x-axis. As with the square stimuli, feature
        and direction assignment were counterbalanced across participants (in
        the displayed example, circles are rotated clockwise up the y-axis.) }
      \label{fig:e3-circles}
    \end{center}
\end{figure}

@@

The consequence of employing a wrapped Gaussian is presented in Figure
\ref{fig:e3-circles}c. Edge effects along the unbounded feature (displayed as
the vertical axis) are no longer present since the edges themselves no longer
exist. Instead, the probability of generating the first Beta exemplar is
bimodally distributed at the edges of the bounded feature (displayed as the
horizontal axis.) Essentially, as with stimuli possessing only bounded features,
the representativeness model still struggles to generate the first Beta
exemplars in the center of the feature space.

It may be argued that our expectation that the representativeness model should
\emph{always} tend to produce first-Betas at the edges may be too strong.
Specifically, this prediction is based on the model that is optimized to data
from Experiments 1 and 2, and it is conceivable that the model may make
drastically different predictions (such as some favoring the initial generation
of centrally-located Betas more than Betas at the edges) in the current
experiment when it is actually fit to empirical data from this experiment.
However, we contend that this prediction of the representativeness model is a
fundamental property of the model due to its assumption of unimodal category
distributions. At the very least, the maximum probability of generating a
centrally located first Beta should almost never be higher than the maximum
probability of generating a first Beta at an edge.

It is also worth noting that since the representativeness model does not explicitly
recognize similarities as distances within the feature space (unlike PACKER),
there is no compression in density regions along the unbounded feature axis.
This provides us with an additional qualitative marker distinguishing the
representativeness model from PACKER -- if the range of the entire Beta category
along the unbounded feature is observed to be shorter than that of the bounded
feature, this would be stronger evidence for PACKER rather than the
representativeness model.

% In the following
% section, we explore -- using both qualitative comparisons and quantitative fits
% -- whether PACKER or the representativeness model is a better account of
% behavioral data in this Corner experiment.

% For the representativeness model, the single unbounded feature essentially
% prevents the formation of edge effects along that feature. Consequently, this results in higher probabilities of the generating the first Beta exemplars along the edges of the bounded feature (Figure \ref{fig:e3-circles}c). As this
% model does not explicitly recognize similarities as distances along the feature
% space (unlike PACKER), it does not experience the compression in density regions
% along the unbounded feature axis. Intuitively, this suggests that the
% representativeness model (which prefers generating exemplars at the boundaries) will prefer to generate stimuli on the boundaries of the bounded feature with some amount of variance over unbounded feature.

@@

We recruited 89 participants from Amazon Mechanical Turk who were randomly
assigned to the bounded, square stimuli ($N=46$), or to the toroidal, circle
stimuli ($N=43$). The square stimuli were exactly the same as in Experiments 1
and 2. The circle stimuli were made up of unfilled circles with a radial line
extending from its center to its edge. These stimuli varied along two features
-- their diameter ranging from 3.0--5.8 cm, and their orientation ranging from
$2.587$--$2\pi + 2.587$ radians, with $0$ radians indicating a completely
horizontal line. The line widths of the circles and radial line were set to 0.05
cm. Similar to Experiments 1 and 2, participants first observed a training phase
where four Alpha exemplars were presented across three blocks followed by the
generation phase, and the features values are divided into 9 discrete steps. The training condition was the Corner category for all
participants.

@@

%Repeat hypotheses
Unlike the previous experiments, Experiment 3 allows us to distinguish between
PACKER and the representativeness model beyond their quantitative fits by
identifying specific qualitatively diagnostic markers. Specifically, if PACKER
is a better account of the data, we would expect a relative decrease in the
range of the entire Beta category along the unbounded feature compared to the
bounded feature, as well as a preference for first-generated Beta exemplars away
from the edges of the feature space. These predictions are correspondingly
reversed if the representativeness model offers a better account of the data.
For our analyses, we define the edge to be the 1-tile wide border of each 9-by-9
large feature space.

Figure \ref{fig:e3-statsboxes} presents the distributional statistics for both
square and circle stimuli. Participants produced circle stimuli with
significantly lower Y-range (lower for the unbounded feature) as compared to 
the square stimuli ($t(87) = 4.94$, $p = .001$), providing early support for PACKER.
There are no significant differences between the distributional statistics of
each generated category between square and circle stimuli (Figure
\ref{fig:e3-statsboxes})

@@

An inspection of the distributions of generated first-Betas (Figure
\ref{fig:e3-behave}) does not immediately indicate support for either model. The
square condition reveals first Betas generated at the corners as well as the
center of the feature space, while the circle condition shows greater
frequencies at the center. However, we also found that 65\% and 87\% of
participants generated their first squares and first circles, respectively, away
from the edge of the feature space.\footnote{Strictly speaking, the circle
  stimuli do not have edges along its orientation feature, However, for ease of
  comparison we will continue to refer to these comparable regions between
  stimuli as edges.} Further, the maximum proportion of first-Betas generated
off the edges (.13) is higher than that of the maximum proportion in an edge
(.04) for the circle stimuli, although the maximum proportions are the same for
squares (.06 for both edge and off-edge.) This preference for generating
first-Betas away from the edge is in line with our hypothesis
supporting PACKER.

\begin{figure}
    \begin{center}
      \includegraphics[width=.6\textwidth]{figs/e3-behave.pdf}
    \caption{Heatmaps indicating the relative frequencies of generating the
      first Beta exemplar in the Corner condition.}
    \label{fig:e3-behave}
    \end{center}
\end{figure}

After fitting the models to data from Experiment 3, there appears to be no major
qualitative deviations in the distribution of first-Betas for the square stimuli
(Figures \ref{fig:e3-firstbetas}a and b) from previous expectations (i.e., as
seen in Figure \ref{fig:e3-squares}), however this is not the case for the
circle stimuli (Figures \ref{fig:e3-firstbetas}c and d compared to Figure
\ref{fig:e3-circles})). Specifically, we see greater variance along the
unbounded feature for PACKER, and most strikingly, that the representativeness
model is able to expand the distribution of generating first-Betas into more
centrally located regions \ref{fig:e3-firstbetas} for circle stimuli. However as
predicted, even with this distribution, the maximum probability of the
representativeness model generating at an edge (.0438) is higher than generating
off-edge (.0394), indicating a preference for producing first-Betas at the edge.
This is in contrast to both PACKER and the empirical data, which both show a
general preference for centrally-located first-Betas.

% These proportions more closely match the predictions of PACKER
% (34\% probability of generating first Beta exemplars at the edges for squares,
% and 24\% probability at edges for circles) compared to the representativeness
% model (53\% of the former, 49\% probability of the latter). Thus, the
% representativeness model's success at capturing Experiments 1 and 2 is in part
% due to the nature of the stimulus features, as well as the structure of
% contrasting category.

\begin{figure}
    \begin{center}
      \includegraphics[width=.6\textwidth]{figs/e3-firstbetas.pdf}
      \caption{Heatmaps indicating the probabilities of each contrast model
        generating the first Beta exemplar in the Corner condition for both
        square and circle stimuli. Models have been fit to data from Experiment 3.}
    \label{fig:e3-firstbetas}
    \end{center}
\end{figure}

The contrast models were fit by optimizing their fit to only Experiment 3. Fit
statistics and parameter values are presented in Table
\ref{table:corner-model-fits}. Although not our main focus in this section, we
also provide the statistics for the classical models -- unsurprisingly, they do
not perform as well as the contrast models. In line with our qualitative
observations, PACKER emerges as the best-fitting model in aggregate.

\begin{table} \centering
  \caption{Results of model-fitting to the combined datasets from Experiments 1,
    2 and 3. Smaller AIC values correspond to better model fits (adjusted
    for number of parameters.)}
\label{table:corner-model-fits}

\begin{tabular}{ l l l l} \\ \\ \textbf{PACKER} & \textbf{Copy \& Tweak} &
\textbf{Representativeness} & \textbf{Hierarchical Bayesian} \\ 
  \hline $AIC = 2954$ &$AIC = 2963$ &$AIC = 2971$  &$AIC = 3025$\\ 
  $L = -1474$          &$L = -1479$   &$L = -1481$    &$L = -1508$\\ 
  $c = 2.08$           &$c = 3.73$    &$\kappa = 0.00$&$\kappa = 0.29$  \\ 
  $\theta_c = 0.87$    &$\theta_c = 0$ (fixed)  &$\nu = 4.07 \times 10^20$ &$\nu
                                                                             = 1.00$\\
  $\theta_t = 2.18$    &$\theta_t = 2.57$       &$\lambda = 0.14$ &$\lambda = 0.00$\\
                       &              &$\theta = 0.73$ &$\theta = 7.50$\\
\end{tabular}
\end{table}

Table \ref{table:corner_counts} presents frequencies of the best-fitting model
for participants in each condition. Overall the representativeness model was
significantly worse at capturing the results of Experiment 3 (it fit 26 out of
the 89 participants better than PACKER; two-tailed Binomial test $p < 0.001$).
Further, this is mostly due to the representativeness model capturing participants
in the unbounded circle condition worse than PACKER (8 of 43 better fit by
representativeness; two-tailed Binomial test, $p<0.001$). The representativeness
model only trended towards fitting participants in the bounded, Square condition
better (18 out of 46; two-tailed Binomial test $p = 0.18$). 
%In addition, these results agree with our prediction that the representativeness model would be poorer in predicting data from circle stimuli (26\% of participants observing circle stimuli) compared to square stimuli (39\% participants observing square stimuli).

\begin{table} \centering
  \caption{Frequencies of individual participants best fit by each contrast
    model  broken down by stimulus type.}
\label{table:corner_counts}
\begin{tabular}{ l l l l l} \\ Stimulus Type & \textbf{PACKER} &
                                                               \textbf{Representativeness}& Total & Two-tailed Binomial test $p$-value \\
  \hline 
  Squares & 28 & 18 & 46 & $p \approx 0.18$\\
  Circles & 35 &  8 & 43 & $p < 0.001$\\
  Total   & 60 & 26 & 89 & $p < 0.001$\\
\end{tabular}
\end{table}

Ultimately, these results suggest that PACKER's exemplar dissimilarity account
of contrast, as opposed to the representativeness account, offers a better
explanation of category generation for the Corner category type. On almost every
measure of analysis in this section, PACKER outperformed the representativeness
model, indicating that the better fit of the latter model demonstrated in
Experiments 1 and 2 is not general to all conditions.

@@

Nonetheless, these classic effects are a core element of the phenomenology of
category generation, and PACKER does not include any mechanisms that explain
them. Instead, through the development and evaluation of the PACKER model, we
have sought to add new elements into such a phenomenology: The broad and strong
influence of category contrast, and the interdependence between category
location and distributional structure. It may be possible to combine the
hierarchical Bayesian approach proposed by \cite{jern2013probabilistic} with
PACKER's underlying claims to obtain a ``best of both worlds'' model, capable of
explaining the role of contrast in category generation, as well as the emulation
of distributional structure. However, as noted in the introduction, the
incorporation of category contrast is antithetical to the core principles of a
traditional, semi-conjugate Bayesian approach. This suggests that category
generation is a fundamentally different computational-level problem
\citep[different from those posed by][]{jern2013probabilistic,kemp2014taxonomy}, which is supported by the representativeness form of their model performing much better.
